\mytitle

### 2016-10-25 paper.tex
/ - finalising work for submission

\documentclass[11pt]{article}
\usepackage[a4paper]{geometry}
\usepackage{clic2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

@@

\newcommand{\mytitle}{bot.zen @ EVALITA 2016 - A minimally-deep learning PoS-tagger \\
    (trained for Italian CMC)}
\newcommand{\mypdftitle}{\mytitle}
\newcommand{\mypdfsubject}{}
\newcommand{\mystitle}{bot.zen - A minimally-deep learning PoS-tagger}
\newcommand{\mykeywords}{PoS tagger, word empbeddings, word2vec, neural
    network, RNN, LSTM, EVALITA 2016, PoSTWITA}
\newcommand{\mypdfkeywords}{\mykeywords}

@@

\begin{abstract} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{English.}  
This article describes the system that participated in the \emph{POS tagging
for Italian Social Media Texts} (PoSTWITA) task of the 5th periodic evaluation
campaign of Natural Language Processing (NLP) and speech tools for the Italian
language EVALITA 2016.

The system continues the work of \newcite{stemle:2016:WAC-X} with minor
modifications and different data sets.  
It combines a small assertion of trending techniques, which implement matured
methods, from NLP and ML to achieve competitive results on PoS tagging of
Italian Twitter texts; in particular, the system uses word embeddings and
character-level representations of word beginnings and endings in a LSTM RNN
architecture.  
Labelled data (Italian UD corpus, DiDi and PoSTWITA) and unlabbelled data
(Italian C4Corpus and PAIS{\`A}) were used for training.

@@

\begin{abstract-alt}
    \textrm{\bf{Italiano.}} Foo Bar
\end{abstract-alt}

%\footnote{\url{http://corpora.ficlit.unibo.it/PoSTWITA/}}
%\footnote{\url{http://www.evalita.it/2016}}

@@

On the other hand, the interest in automatic evaluation of social media texts,
in particular for microblogging texts such as tweets, has been growing
considerably, and specialised tools for Twitter data have become available for
different languages. 
But Italian completely lacks such resources, both regarding annotated corpora
and specific PoS-tagging tools.\footnote{http://www.evalita.it/2016/tasks/postwita}
To this end, the \emph{POS tagging for Italian Social Media Texts} (PoSTWITA)
task was proposed for EVALITA 2016 concerning the domain adaptation of
PoS-taggers to Twitter texts.
The systems were evaluated by the organisers on unlabelled but pre-tokenised
data (4759 tokens in 301 tweets).

@@

The paper is organised as follows: We present our system design in
Section~\ref{sec:design}, the implementation in
Section~\ref{sec:implementation}, and its evaluation in
Section~\ref{sec:casestudies}. 
Section~\ref{sec:conclusion} concludes with an outlook on possible
implementation improvements.

@@

Overall, our design takes inspiration from as far back as~\newcite{Benello1989}
who used four preceding words and one following word in a feed-forward neural
network with backpropagation for PoS tagging, builds upon the strong foundation
laid down by~\newcite{collobert:2011b} for a NN architecture and learning
algorithm that can be applied to various natural language processing tasks, and
ultimately is a variation of
% achieved an accuracy of $0.95$ on the Brown Corpus.
~\newcite{santos2014learning} who trained a NN for PoS tagging, with 
character-level and WE representations of words. 
%On the Wall Street Journal (WSJ) portion of the Penn Treebank they achieve
%accuracy scores of $0.9732$.

@@

A straight-forward approach to incorporate WEs into feature-based
language models is to use the embeddings' vector representations as
features.\footnote{For an overview see,
e.g.~\newcite{Turian:2010:WRS:1858681.1858721}.}
Having said that, WEs are also used in neural network architectures, where they
constitute (part of) the input to the network.

Neural networks (NNs) consist of a large number of simple, highly interconnected
processing nodes in an architecture loosely inspired by the structure of the
cerebral cortex of the brain~\cite{oreilly2000}.
The nodes receive weighted inputs through these connections and \emph{fire}
according to their individual thresholds of their shared activation function.
A firing node passes on an activation to all successive connected nodes.
During learning the input is propagated through the network and the output is
compared to the desired output. 
Then, the weights of the connections (and the thresholds) are adjusted
step-wise so as to more closely resemble a configuration that would produce the
desired output.
After all input cases have been presented, the process typically starts over
again, and the output values will usually be closer to the correct values.

@@

Overall, with this design we not only benefit from available labelled data but
also from available general and domain-specific unlabelled data.

@@

We maintain the implementation in a source code repository at
\url{https://github.com/bot-zen/}.  
The version tagged as {\tt 1.1} comprises the version that was used to generate
the results submitted to the ST.

@@

%When computing WEs we keep in mind that \newcite{TACL570} observed that one
%specific configuration of~\wtv, namely the skip-gram model with negative
%sampling (SGNS) 
%"is a robust baseline.  While it might not be the best method for every task,
%it does not significantly underperform in any scenario. Moreover, SGNS is the
%fastest method to train, and cheapest (by far) in terms of disk space and
%memory consumption".
%Coincidentally,~\newcite{arXiv:1310.4546} also suggest to use SGNS
We incorporates \wtv's original C implementation for learning WEs%
\footnote{\url{https://code.google.com/archive/p/word2vec/}}
in an independent pre-processing step, i.e.~we pre-compute the WEs.
Then, we use gensim%
\footnote{\url{https://radimrehurek.com/gensim/}}%
, a Python tool for unsupervised semantic modelling from plain text, to load
the data, and to extract the vector representations of the embedded words as
input to our NN.

@@

Our implementation uses Keras, a minimalist, highly modular NNs library,
written in Python and capable of running on top of either TensorFlow or
Theano~\cite{chollet2015}. 
In our case it runs on top of Theano, a Python library that allows to define,
optimize, and evaluate mathematical expressions involving multi-dimensional
arrays efficiently~\cite{Theano2016}.

The input to our network are sequences of the same length as the sentences we
process.
During training we group sentences of the same length into batches.
Each single word in the sequence is represented by its sub-word information and
two WEs that come from two sources (see Section~\ref{sec:casestudies}).
%Note that the WEs of the domain-specific corpus may have only been derived
%from a small corpus.
Unknown words, i.e.~words without a WE, are mapped to a randomly generated
vector representation once, and this representation is reused later.
In Total, each word is represented by $1,800$ features: two times $500$ (WEs),
and ten times $80$ for two 5-grams (word beginning and ending).
(If words are shorter than 5 characters their 5-grams are zero-padded.)

This sequential input is fed into a LSTM layer that, in turn, projects to a
fully connected output layer with softmax activation function.
We use categorical cross-entropy as loss function and backpropagation in
conjunction with the RMSprop optimization for learning.  
At the time of writing, this was the Keras default---or the explicitly
documented option to be used---for our type of architecture.

\section{Case Study} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:casestudies}

% For the proposed task we plan to re-use the tweets being part of the EVALITA2014 SENTIPLOC corpus: both the development and test set will be manually annotated for a global amount of 4041+1749 tweets and will be distributed as the new development set. Then a new manually annotated test set composed of 600/700 tweets will be produced using texts from the same period of time. All the annotations will be carried out by three different annotators. 
% 
% We will distribute a tokenised version of the texts in order to avoid tokenisation problems among participants and the boring problem of disappeared tweets.

We used our implementation to participate in the EmpiriST 2015 shared task.
First, we describe the corpora used for training, and then the specific system
configuration(s) for the ST.
%% and subsequently, to run experiments with different data and parameters for
%%training (see Subsection~\ref{ssec:case_postempirist}).
%Following, a short description of the data we used for training.

\subsection{Training Data for \wtv and PoS Tagging} %%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Tiger v2.2 (PoS)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{Tiger v2.2}%
\footnote{\url{http://www.ims.uni-stuttgart.de/forschung/ressourcen/korpora/tiger.html}}
is version 2.2 of the TIGERCorpus~\cite{Brants2004} containing German newspaper
texts. 
The corpus was semi-automatically PoS tagged, and is one of the standard
corpora used for German PoS tagging. 
It contains 888,238 tokens in 50,472 sentences.
For research and evaluation purposes, the TIGERCorpus can be downloaded for
free.

\subsubsection{German Wikipedia (\wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{de.wiki'15}%
\footnote{\url{http://www1.ids-mannheim.de/kl/projekte/korpora/verfuegbarkeit.html\#Download}}
are user talk pages (messages from users to users, often questions and advice),
article talk pages (questions, concerns or comments related to improving a
Wikipedia article), and article pages of the German wikipedia from 2015, made
available by the \emph{Institut f\"{u}r Deutsche
Sprache}\footnote{\url{http://www.ids-mannheim.de}}.
%% user talk pages
%% 378,576,711 toks
%% 15,343,904 s
%%
%% article talk pages
%% 446,551,142 toks 
%% 17,275,385 s
%%
%% article pages
%% 1,126,462,560 toks
%% 46,740,611 s
%%
% tokens: 378,576,711+446,551,142+1,126,462,560 = 1,951,590,413
% sentences: 15,343,904+17,275,385+46,740,611 = 79,359,900
The corpus contains 2 billion tokens 
(talk:379m, article talk:447m, article:1,1bn) 
in 79 million sentences 
(talk:15m, article talk:17m, article:47m), is well-sized for \wtv, and
also (partly) resembles or target data.
It is available under the CC BY-SA 3.0\footnote{Creative Commons
Attribution-ShareAlike 3.0 Unported, i.e.~the data can be copied and
redistributed, and adapted for any purpose, even commercially. See
\url{http://creativecommons.org/licenses/by-sa/3.0/} for more details.}
license.

\subsubsection{EmpiriST 2015 Data (PoS and \wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{subsec:empirist}
\emph{empirist}%
\footnote{\url{https://sites.google.com/site/empirist2015/home/shared-task-data}}
is the CMC and Web data made available by the organizers of the ST.
It contains data samples from different CMC genres and samples from text genres
on the Web.
The training corpus contains 10,053 tokens and was PoS tagged by two annotators
(unclear cases were decided by a third person). 
The trial corpus contains around 3,600 tokens (2,100 CMC%
\footnote{For evaluation during the development phase we used
    \emph{empirist}-trial.
    Unfortunately, we found out only later that the CMC part of the trial data
    is also part of the training data, i.e.~for the CMC data our evaluation
    data was identical with the training data.}%
, 1,500 Web) and was PoS tagged by one annotator (without systematic error
checks%
%, i.e.~the accuracy of the PoS annotations is likely lower than on the
%training data and on the ST final evaluation data
).
See~\newcite{empirist2016} for more details.

@@

%\begin{table}[h]
%\begin{center}
%\begin{tabular}{l|r|r}
%\hline
%\multicolumn{1}{c}{} & \multicolumn{1}{c}{Tokens} & \multicolumn{1}{c}{Sentences} \\ \hline
%empirist       & $10*10^3$     & $0.6*10^3$    \\ \hline
%Tiger v2.2     & $0.9*10^9$    & $0.05*10^6$   \\ \hline\hline
%de.wiki'15   	& $2.0*10^9$    & $79*10^6$     \\ \hline
%%C4Corpus      & $$            & $$            \\ \hline
%%DECOW14       & $11.7*10^9$   & $600*10^6$    \\ \hline
%%deTenTen	    & $19.9*10^9$   & $1300*10^6$   \\ \hline
%\end{tabular}
%\end{center}
%\caption{\label{tab:corpora}Training data used for the EmpiriST 2015 shared
%    task. The PoS tags of the upper two corpora were used for training.
%    % (upper part) and additional data for the subsequent experiments (lower
%    % part).
%} \end{table}

%\emph{C4Corpus}%
%\footnote{\url{https://github.com/dkpro/dkpro-c4corpus}} is a full documents
%German Web corpus that has been extracted from CommonCrawl, the largest
%publicly available general Web crawl to date. 
%The data is available under the CreativeCommons license
%family\footnote{\url{https://creativecommons.org/licenses/}}. 
%See Habernal~\shortcite{Habernal.et.al.2016.LREC} for details about the corpus
%construction pipeline, and other information about the corpus.
%
%\emph{DECOW14}%
%\footnote{\url{http://corporafromtheweb.org/decow14/}} is the sentence shuffled
%German Web corpus by COW created with the 2014 technology of the COW
%initiative~\cite{SchaeferBildhauer2012}.
%The data can be used for purposes of academic research.
%% $11.7*10^9$ tokens 
%% $0.6*10^9$ (624,767,747) sentences
%
%\emph{deTenTen} Web corpus.
%%DeTenTen
%% $19.9*10^9$ (19,918,263,493) tokens
%% $1.3*10^9$ (1,309,696,220) sentences
%\cite{1120431}

\subsection{EmpiriST 2015 shared task} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{ssec:case_empirist}

For the ST we used one overall configuration for the system, but we used three
different corpus configurations for training.
Consequently, we participated in the ST with three runs:
we used PoS tags from \emph{empirist} (run 1), from \emph{Tiger v2.2} (run 2),
and from both (run 3).  
For \wtv we trained a 500-dimensional skip-gram model on \emph{empirist} that
ignored all words with less than 3 occurrences within a window size of 10;
it was trained with negative sampling (value 5) and erroneously%
\footnote{%
    According to \wtv's author, technically negative sampling and
    hierarchical softmax can be combined but one should avoid combining them~%
    (see~\url{https://groups.google.com/forum/\#!topic/word2vec-toolkit/WUWad9fL0jU}).

    We had forgotten to deactivate an option in a data processing script.
}
also with hierarchical softmax.
We also trained a 500-dimensional continuous bag-of-words model on
\emph{de.wiki'15} that ignored all words with less than 25 occurrences within a
window size of 10; it was trained with negative sampling (value 3) and
erroneously also with hierarchical softmax.

The rational behind training the two models differently was that according to
\wtv author's experience%
\footnote{\url{https://groups.google.com/d/msg/word2vec-toolkit/NLvYXU99cAM/E5ld8LcDxlAJ}}
a skip-gram model "works well with small amount[s] of the training data, [and]
represents well even rare words or phrases", and a cbow model is "several times
faster to train than the skip-gram, [and has] slightly better accuracy for the
frequent words".
The other \wtv parameters were left at their default settings%
\footnote{\texttt{-sample 1e-3 -iter 5 -alpha 0.025} for skip-gram and
\texttt{-alpha 0.05} for continuous bag-of-words}.
% empirist.bin -cbow 0 -size 500 -window 10 -hs 1 -negative 5 -min-count 3 
% bigdata.bin -cbow 1 -size 500 -window 10 -hs 1 -negative 3 -min-count 25

To optimize the system's output we ran a simple grid search for three
parameters: the hidden LSTM layer's size, the dropout value for the projections
from the LSTM to the output layer during training, and the number of epochs
during training. 
The found values were size:$1024$, dropout:$0.1$, epochs:$20$.
%lstm_output_dims = [128, 256, 512, 1024*]
%dropouts = [0.1*, 0.25, 0.5, 0.75]
%nb_epochs = [5, 10, 20*]

\begin{table}[h]
\begin{center}
\begin{tabular}{l|r|r}
\hline
\multicolumn{1}{c}{} & \multicolumn{1}{c}{CMC} & \multicolumn{1}{c}{Web} \\ \hline
(1) \emph{empirist}             & $81.03$       & $86.97$ \\ \hline
(2) \emph{Tiger v2.2}           & $73.56$       & $89.73$ \\ \hline
(3) \emph{empirist}+\emph{Tiger v2.2} & $85.42$ & $90.63$ \\ \hline\hline
Winning Team                    & $87.33$       & $93.55$ \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:results}Official results of our PoS tagger for the three
runs on the EmpiriST 2015 shared task data.
} 
\end{table}

%The official results are listed in Table~\ref{tab:results}.

% 1: empirist_training (tagset:ibk)
% 2: tiger-2.2 (tagset:1999)
% 3: empirist_training + tiger-2.2 (tagset:ibk)
%Our evaluation on the EmpiriST trial data resulted in these accuracy scores:
%1:~$0.88$, 
%2:~$0.78$, 
%3:~$0.93$;
%and the overall ST evaluation was:
%1:~$0.84$,
%2:~$0.82$,
%3:~$0.88$.

%\subsection{Post-EmpiriST Experiments}
%\label{ssec:case_postempirist}
%As a post-EmpiriST addition to the results we wanted to (1)~remedy some errors
%and oversights in our initial pre-processing and training, (2)~optimize the
%learning parameters of the word embedding models and the neural network, and
%(3)~train word embedding models on other data.
%\fxnote[margin]{add (1)}
%
%The optimal learning parameters for the word embedding models were selected
%according to a comprehensive set of experiments and comparisons by
%Levy~\shortcite{TACL570}, and observations and suggestions by
%Mikolov~\shortcite{arXiv:1310.4546}. 
%For the post-EmpiriST experiments we used the skip-gram with negative-sampling
%training method (SGNS), with a negative-sampling value of $15$ (and without
%hierarchical softmax, cf.~Subsection~\ref{ssec:case_empirist}), ignored all
%words with less than $100$ occurrences within a window size of $10$, and a
%sub-sampling value of $10^{-5}$. 
%% cbow=0, negative=15, hs=0, window=10, sample=10^-5 
%% arXiv:1310.4546: -negative 15 -sample 0.00001
%% TACL570: alpha=0.75 and dyn=1-> fixed in word2vec algorithm;
%%There are quite a few differences between the skip-gram and the CBOW models.
%%However, if you have a lot of training data, their performance should be
%%comparable.%
%%\footnote{\url{https://groups.google.com/d/msg/word2vec-toolkit/NLvYXU99cAM/E5ld8LcDxlAJ}}
%
%As additional data to train word embedding models we obtained and processed
%\emph{C4Corpus}, \emph{DECOW14}, and \emph{deTenTen}.
%
%\fixme[inline]{The system is said to be trained on various widely available
%    corpora of German such as Tiger 2.2, DECOW148 etc. It is not discussed in
%    any detail and thus does not become clear why these corpora were selected
%    and especially what the systematic benefit of using these different data
%    sets was; more specifically, the paper does not systematically explore the
%    diversity of the data and their contribution to a learning improvement on
%    the system. A section with a systematic evaluation is missing and so is a
%    section with a more thorough discussion as to why this particular approach
%    was selected. It is suggested to the authors to provide more background on
%    these questions.}

@@

We presented our submission to the EmpiriST 2015 shared task, where we
participated in the PoS tagging sub-task with fair results on the CMC data and
adequate results on the Web data. 
Still, our implementation, albeit following state-of-the art designs and methods, is
quite unpolished, and can certainly gain performance with more detailed tuning.
For example, adding special sequence start and sequence stop symbols to the
input is typically done as a pre-processing step, which might improve the
results at the beginning and the end of sentences;
or we might gain some performance by adding additional hidden layers to enable
the network to learn more intermediate abstractions.
A more profound design change could also help, e.g.~Recurrent Memory Network
are a novel recurrent architecture that have been shown to outperform LSTMs on
some language modelling tasks.
%and sentence completion tasks, while also allowing insights into the data from
%a linguistic perspective~\newcite{tran-bisazza-monz:2016:N16-1}.
Finally, for learning the word embeddings we could use different corpora, or
selectively extract parts from large web-corpora resembling---as much as
possible---the type of data that is to be tagged.

>>>>

\documentclass[11pt]{article}
\usepackage[a4paper]{geometry}
\usepackage{clic2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{textcomp}

@@

\newcommand{\mytitle}{bot.zen @ EVALITA 2016 - A minimally-deep learning PoS-tagger \\
    (trained for Italian Tweets)}
\newcommand{\mypdftitle}{\mytitle}
\newcommand{\mypdfsubject}{}
\newcommand{\mystitle}{bot.zen - A minimally-deep learning PoS-tagger}
\newcommand{\mykeywords}{PoS tagger, word empbeddings, word2vec, neural
    network, RNN, LSTM, EVALITA 2016, PoSTWITA}
\newcommand{\mypdfkeywords}{\mykeywords}

@@

\begin{abstract} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{English.}  
This article describes the system that participated in the \emph{POS tagging
for Italian Social Media Texts} (PoSTWITA) task of the 5\textsuperscript{th}
periodic evaluation campaign of Natural Language Processing (NLP) and speech
tools for the Italian language EVALITA 2016.

The work is a continuation of \newcite{stemle:2016:WAC-X} with minor
modifications to the system and different data sets.  
It combines a small assertion of trending techniques, which implement matured
methods, from NLP and ML to achieve competitive results on PoS tagging of
Italian Twitter texts; in particular, the system uses word embeddings and
character-level representations of word beginnings and endings in a LSTM RNN
architecture.  
Labelled data (Italian UD corpus, DiDi and PoSTWITA) and unlabbelled data
(Italian C4Corpus and PAIS{\`A}) were used for training.

@@

\begin{abstract-alt}
\textrm{\bf{Italiano.}}
Questo articolo viene descritto il sistema che ha partecipato alla \emph{POS
tagging for Italian Social Media Texts} (PoSTWITA) task del
5\textdegree~campagna di valutazione periodica degli strumenti del Natural
Language Processing (NLP) e gli strumenti del linguaggio per la lingua italiana
EVALITA 2016.

Il sistema continua l'opera di \newcite{stemle:2016:WAC-X} con piccole
modifiche e diversi set di dati. 
Esso combina una piccola affermazione di tecniche di tendenza, che implementano
maturati metodi, dalla NLP e ML di raggiungere risultati competitivi in pos
codifica dei testi Twitter italiani; in particolare, il sistema utilizza
incastri di parole e rappresentazioni a livello di carattere di inizio di
parole e terminazioni in un'architettura LSTM RNN. 
Dati etichettati (Italiano corpus UD, Didi e PoSTWITA) ei dati non etichettati
(italiano C4Corpus e Paisà) sono stati utilizzati per la formazione.

Il sistema è disponibile sotto la licenza open-source APLv2.
\end{abstract-alt}

@@

On the other hand, the interest in automatic evaluation of social media texts,
in particular for microblogging texts such as tweets, has been growing
considerably, and specialised tools for Twitter data have become available for
different languages. 
But Italian completely lacks such resources, both regarding annotated corpora
and specific PoS-tagging tools.\footnote{http://www.evalita.it/2016/tasks/postwita}
To this end, the \emph{POS tagging for Italian Social Media Texts} (PoSTWITA)
task was proposed for EVALITA 2016 concerning the domain adaptation of
PoS-taggers to Twitter texts.

@@

The paper is organised as follows: We present our system design in
Section~\ref{sec:design}, the implementation in
Section~\ref{sec:implementation}, and its evaluation in
Section~\ref{sec:results}. 
Section~\ref{sec:conclusion} concludes with an outlook on possible
implementation improvements.

@@

Overall, our design takes inspiration from as far back as~\newcite{Benello1989}
who used four preceding words and one following word in a feed-forward neural
network with backpropagation for PoS tagging, builds upon the strong foundation
laid down by~\newcite{collobert:2011b} for a neural network (NN) architecture
and learning algorithm that can be applied to various natural language
processing tasks, and
ultimately is a variation of
% achieved an accuracy of $0.95$ on the Brown Corpus.
~\newcite{santos2014learning} who trained a NN for PoS tagging, with 
character-level and WE representations of words. 
%On the Wall Street Journal (WSJ) portion of the Penn Treebank they achieve
%accuracy scores of $0.9732$.

@@

A straight-forward approach to incorporate WEs into feature-based
language models is to use the embeddings' vector representations as
features.\footnote{For an overview see,
e.g.~\newcite{Turian:2010:WRS:1858681.1858721}.}
Having said that, WEs are also used in NN architectures, where they constitute
(part of) the input to the network.

Neural networks consist of a large number of simple, highly interconnected
processing nodes in an architecture loosely inspired by the structure of the
cerebral cortex of the brain~\cite{oreilly2000}.
The nodes receive weighted inputs through these connections and \emph{fire}
according to their individual thresholds of their shared activation function.
A firing node passes on an activation to all successive connected nodes.
During learning the input is propagated through the network and the output is
compared to the desired output. 
Then, the weights of the connections (and the thresholds) are adjusted
step-wise so as to more closely resemble a configuration that would produce the
desired output.
After all input cases have been presented, the process typically starts over
again, and the output values will usually be closer to the correct values.

@@

Overall, with this design we not only benefit from available labelled data but
also from available general or domain-specific unlabelled data.

@@

We maintain the implementation in a source code repository at
\url{https://github.com/bot-zen/}.  
The version tagged as {\tt 1.1} comprises the version that was used to generate
the results submitted to the shared task (ST).

@@

When computing WEs we take into consideration \newcite{TACL570}: they observed
that one specific configuration of~\wtv, namely the skip-gram model with
negative sampling (SGNS) 
"is a robust baseline.  While it might not be the best method for every task,
it does not significantly underperform in any scenario. Moreover, SGNS is the
fastest method to train, and cheapest (by far) in terms of disk space and
memory consumption".
Coincidentally,~\newcite{arXiv:1310.4546} also suggest to use SGNS.
We incorporate \wtv's original C implementation for learning WEs%
\footnote{\url{https://code.google.com/archive/p/word2vec/}}
in an independent pre-processing step, i.e.~we pre-compute the WEs.
Then, we use gensim%
\footnote{\url{https://radimrehurek.com/gensim/}}%
, a Python tool for unsupervised semantic modelling from plain text, to load
the pre-computed data, and to compute the vector representations of
input words for our NN.

@@

Our implementation uses Keras, a high-level NNs library, written in Python and
capable of running on top of either TensorFlow or Theano~\cite{chollet2015}. 
In our case it runs on top of Theano, a Python library that allows to define,
optimize, and evaluate mathematical expressions involving multi-dimensional
arrays efficiently~\cite{Theano2016}.

The input to our network are sequences of the same length as the sentences we
process.
During training, we group sentences of the same length into batches and process
the batches according to sentence length in increasing order.
Each single word in the sequence is represented by its sub-word information and
two WEs that come from two sources (see Section~\ref{sec:results}).
%Note that the WEs of the domain-specific corpus may have only been derived
%from a small corpus.
For unknown words, i.e.~words without a pre-computed WE, we first try to
find the most similar WE considering 10 surrounding words.  
If this fails, the unknown word is mapped to a randomly generated vector
representation.
In Total, each word is represented by $2,280$ features: two times $500$ (WEs),
and sixteen times $80$ for two 8-grams (word beginning and ending).
If words are shorter than 8 characters their 8-grams are zero-padded.

This sequential input is fed into a LSTM layer that, in turn, projects to a
fully connected output layer with softmax activation function.
During training we use dropout for the projection into the output layer,
i.e.~we set a fraction ($0.5$) of the input units to 0 at each update, which
helps prevent overfitting~\cite{Srivastava2014}.
We use categorical cross-entropy as loss function and backpropagation in
conjunction with the RMSprop optimization for learning.  
At the time of writing, this was the Keras default---or the explicitly
documented option to be used---for our type of architecture.

\section{Results} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:results}

We used our slightly modified implementation to participate in the \emph{POS
tagging for Italian Social Media Texts} (PoSTWITA) shared task (ST) of the
5\textsuperscript{th} periodic evaluation campaign of Natural Language
Processing (NLP) and speech tools for the Italian language EVALITA 2016.
First, we describe the corpora used for training, and then the specific system
configuration(s) for the ST.

\subsection{Training Data for \wtv and PoS Tagging} %%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{DiDi-IT (PoS, \wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{didi-it}~\cite{frey-glaznieks-stemle:2016:didi}
(version September 2016) is the Italian sub-part of the DiDi corpus, a corpus
of South Tyrolean German and Italian from Facebook (FB) users' wall posts,
comments on wall posts and private messages.

The Italian part consists of around 100,000 tokens collected from 20 profiles
of Facebook users residing in South Tyrol.
This version has about 20,000 PoS tags semi-automatically corrected by a single
annotator.

The anonymised corpus is freely available for research purposes.

\subsubsection{Italian UD (PoS, \wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Universal Dependencies (UD) is a project that is developing
cross-linguistically consistent treebank annotation for many
languages.\footnote{\url{http://universaldependencies.org/}}

\emph{italian-UD}%
\footnote{\url{http://universaldependencies.org/it/overview/introduction.html}}
(version from January 2015) corpus was originally obtained by conversion from
ISDT (Italian Stanford Dependency Treebank) and released for the dependency
parsing ST of EVALITA 2014~\cite{Bosco2014}.
The corpus has semi-automatically converted PoS tags from the original two
Italian treebanks, differing both in corpus composition and adopted annotation
schemes.

The corpus contains around 317,000 tokens in around 13,000 sentences from
different sources and genres.  
It is available under the CC BY-NC-SA 3.0\footnote{Creative Commons
Attribution-NonCommercial-ShareAlike 3.0 Unported, i.e.~the data can be copied
and redistributed, and adapted for purposes other than commercial ones. See
\url{https://creativecommons.org/licenses/by-nc-sa/3.0/} for more details.}
license.

\subsubsection{PoSTWITA (PoS and \wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sssec:postwita}
\emph{postwita} is the Twitter data made available by the organizers of the ST.
It contains Twitter tweets from the EVALITA2014 SENTIPLOC corpus: the
development and test set and additional tweets from the same period of time
were manually annotated for a global amount of 6438 tweets (114,967 tokens) and
were distributed as the development set. 
The data is PoS tagged according to UD but with the additional insertion of
seven Twitter-specific tags.
All the annotations were carried out by three different annotators.
The data was only distributed to the task participants.

\subsubsection{C4Corpus (\wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{c4corpus}%
\footnote{\url{https://github.com/dkpro/dkpro-c4corpus}} is a full documents
Italian Web corpus that has been extracted from CommonCrawl, the largest
publicly available general Web crawl to date. 
See Habernal~\shortcite{Habernal.et.al.2016.LREC} for details about the corpus
construction pipeline, and other information about the corpus.

The corpus contains about $670m$ tokens in $22m$ sentences.
The data is available under the CreativeCommons license
family.

\subsubsection{PAIS{\`A} (\wtv)} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\emph{paisa}~\cite{paisa2014} is a corpus of authentic contemporary Italian
texts from the web (harvested in September/October 2010).
It was created in the context of the project PAIS{\`A} (P{\'i}attaforma per
l'Apprendimento dell'Italiano Su corpora Annotati) with the aim to provide a
large resource of freely available Italian texts for language learning by
studying authentic text materials.

The corpus contains about $270m$ tokens in about $8m$ sentences. 
The data is available under the CC BY-NC-SA
3.0\footnote{\url{https://creativecommons.org/licenses/by-nc-sa/3.0/}} license.

@@

%\begin{table}[h]
%\begin{center}
%\begin{tabular}{l|r|r}
%\hline
%\multicolumn{1}{c}{} & \multicolumn{1}{c}{Tokens} & \multicolumn{1}{c}{Sentences} \\ \hline
%corpus 1       & $10*10^3$     & $0.6*10^3$    \\ \hline
%corpus 2       & $0.9*10^9$    & $0.05*10^6$   \\ \hline\hline
%corpus 3     	& $2.0*10^9$    & $79*10^6$     \\ \hline
%%C4Corpus      & $$            & $$            \\ \hline
%\end{tabular}
%\end{center}
%\caption{\label{tab:corpora}Training data used for the shared
%    task. The PoS tags of the upper two corpora were used for training.
%} \end{table}

\subsection{PoSTWITA shared task} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{ssec:postwita}

For the ST we used one overall configuration for the system but three different
corpus configurations for training.
However, only one corpus configuration was entered into the ST: we used PoS
tags from \emph{didi-it + postwita} (run 1), from \emph{italian-UD} (run 2),
and from both (run 3).
For \wtv we trained a 500-dimensional skip-gram model on \emph{didi-it +
italian-UD + postwita} that ignored all words with less than 2 occurrences
within a window size of 10; it was trained with negative sampling (value 15).
We also trained a 500-dimensional skip-gram model on \emph{c4corpus + paisa}
that ignored all words with less than 33 occurrences within a window size of
10; it was trained with negative sampling (value 15).

The other \wtv parameters were left at their default settings%
\footnote{\texttt{-sample 1e-3 -iter 5 -alpha 0.025}}.

The evaluation of the systems was done by the organisers on unlabelled but
pre-tokenised data (4759 tokens in 301 tweets), and was based on a
token-by-token comparison.
The considered metric was accuracy, i.e.~the number of correctly assigned PoS
tags divided by the total number of tokens.

\begin{table}[h]
\begin{center}
\begin{tabular}{l|r|r}
\hline
\textbf{(1)} \textbf{\emph{didi-it + postwita}}   & $\textbf{76.00}$        \\ \hline
(2) \emph{italian-UD}           & $80.54$        \\ \hline
(3) \emph{didi-it + postwita + italian-UD} & $81.61$  \\ \hline\hline
Winning Team                    & $93.19$        \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:results}Official result(s) of our PoS tagger for the
three runs on the PoSTWITA ST data.
} 
\end{table}

We believe, the unexpectedly little performance gain from utilizing the much
larger \emph{italian-UD} data over the rather small \emph{didi-it + postwita}
data may be rooted in the insertion of Twitter-specific tags into the data
(see~\ref{sssec:postwita}), something we did not account for, i.e.~$18,213$ of
$289,416$ and more importantly $7,778$ of $12,677$ sentences had imperfect
information during training.

%\fixme[inline]{The system is said to be trained on various widely available
%    corpora of German such as Tiger 2.2, DECOW148 etc. It is not discussed in
%    any detail and thus does not become clear why these corpora were selected
%    and especially what the systematic benefit of using these different data
%    sets was; more specifically, the paper does not systematically explore the
%    diversity of the data and their contribution to a learning improvement on
%    the system. A section with a systematic evaluation is missing and so is a
%    section with a more thorough discussion as to why this particular approach
%    was selected. It is suggested to the authors to provide more background on
%    these questions.}

@@

We presented our submission to the PoSTWITA ST, where we participated
with moderate results. 
In the future, we will try to rerun the experiment with training data that
takes into consideration the Twitter-specific tags of the ST.
